---
title: "R Notebook"
output: html_notebook
---

## 2.6 Practice

### Easy

1. 2 - Pr(rain|Monday)
2. 3 - The probability of it being monday, given it is raining
3. 1 - Pr(Monday|rain)
4. If you were to toss the globe many times the proportion of times you would get water is 0.7.

### Medium

1.1 W, W, W

```{r}
grid = seq(0, 1, 0.05)
prior = rep(1, length(grid)) / length(grid)

# The likelihood of 3 waters from 3 tosses
lik = dbinom(x = 3, size = 3, prob = grid)

posterior = prior * lik
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```

1.2 W, W, W, L

```{r}
lik = dbinom(x = 3, size = 4, prob = grid)
posterior = prior * lik
head(posterior)
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```

1.3 L, W, W, L, W, W, W

Seven trials, five waters

```{r}
lik = dbinom(x = 5, size = 7, prob = grid)
posterior = prior * lik
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```

2.1 Same data, but prior=0 when p < 0.5.

```{r}
over_half = grid >= 0.5
prior = as.numeric(over_half) / sum(over_half)
plot(grid, prior, type = "b")
```

3 waters from 3 trials

```{r}
posterior = prior * dbinom(3, 3, grid)
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```

2.2 W, W, W, L

```{r}
posterior = prior * dbinom(3, 4, grid)
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```

2.3 L, W, W, L, W, W, W

```{r}
posterior = prior * dbinom(5, 7, grid)
posterior = posterior / sum(posterior)
plot(grid, posterior, type = "b")
```


2.3 What is the probability of the globe being earth, given seen land

```{r}
# P(Earth)
prior = 0.5

# Likelihood, P( Land | Earth )
# 70% water = 0.3 land
lik = 0.3

# P( Land )  (How surprised are you at seeing land?)
p_land = (0.5 * 1) + (0.5 * 0.3)

posterior = (prior * lik) / p_land
posterior
```

2.4 Deck with three cards

- WW - two white sides
- BB - two black sides
- BW - 1 black, 1 white side

Cards are shuffled, 1 drawn at random. Shows a black side. What is the probability the other side is black?

Could be any of the B sides. 2/3 of the time it will be BB. Therefore the other side is B. 1/3 of the time it is BW, therefore the other side is W. 

- WW could have produced the data 0 ways.
- BB could have produced the data 2 ways.
- BW could have produced the data 1 way.

Therefore of all the (2 + 1) = 3 ways to draw a black card. 2/3 of the time it is BB and therefore 2/3 the time the other side is B.

2.5 Deck with four cards

- 2x BB
- 1x BW
- 1x WW

Probability the other side is black = 4/5.

Now, four out of the five possible ways to generate a black card result in the other side also being black.

2.6  Heavy black ink. Less likely to pull a black card.

Back to three cards:

- BB
- BW
- WW

BB : BW : WW pulls are in ratio of 1 : 2 : 3.

Card is pulled and black side appears. What is the probability the other side is black?

Ways to pull a black card:

- BB, 2 ways, 1 weight -> 2
- BW, 1 way, 2 weight -> 2
- WW, 0 ways, 3 weight -> 0

Now 2/4 = 0.5 of the time the second face will be black.

2.7 2 card draws

Original setup (2.4).

- Draw 1 is black.
- Draw 2 is white.

What is the probability the other side of draw 1 is black?

These are the ways cards could be drawn, where draw 1 is black and draw 2 is white:

- BB, BW (x2)
- BB, WW (x2, x2)
- BW, WW (x2)

(Within BB there are 2 ways of drawing B first, (same for WW), so these counts are doubled.)

(All other sequences have 0 likelihood:)

- BW, BB
- WW, BB
- WW, BW

Of the possible ways 6 / 8 = 3 / 4 occur with drawing the BB card first.

### Hard

1. Pandas

T = twins
A = species A
B = species B

P(T|A) = 0.1
P(T|B) = 0.2

a panda of unknown species has given birth to twins. what is the probability the next birth will be twins?

data = T

1/3 times it will be A. multiply by 0.1 to get p(Twins)

2/3 times it will be B. multiply by 0.2 to get p(Twins)

```{r}
1/3 * 0.1 + 2/3 * 0.2
```

2. Pandas again

Compute probability panda is species A, given we've observed one birth of twins.

P(A|T) = P(T|A)P(A) / P(T)

```{r}
(0.1 * 0.5) / (0.5 * 0.1 + 0.5 * 0.2)
```

3. Pandas again again

Same panda in question above gives birth again and it is not twins.  Compute posterior probability that this panda is species A.

Now, P(A) is 1/3 (computed in previous question), not 0.5.

- S = singleton
- P(A|S) = P(S|A)P(A) / P(S)
- P(A) = 1/3
- P(S|A) = 0.9
- P(S|B) = 0.8
- P(S) = P(A) * P(S|A) + P(B) * P(S|B)

```{r}
(0.9 * (1/3)) / ((1/3) * 0.9 + (2/3) * 0.8)
```

4. Genetic test

- Probability correctly identifies species A is 0.8.
- Probability correctly identifies species B is 0.65.

Vet administers test, which is positive for A. What is the posterior probability of species A, given no other information.

```{r}
# prior P(A)
p_a = 0.5

# likelihood P(Test=A|A)
lik = 0.8

# data
p_data = p_a * 0.8 + (1 - p_a) * (1-0.65)

post = p_a * lik / p_data
post
```

What is the posterior probability of species A, given the birth information?

```{r}
# now prior, P(A) is 0.36 (computed in Q3.)
p_a = 0.36
p_data = p_a * 0.8 + (1 - p_a) * (1-0.65)
post = p_a * lik / p_data
post
```


