---
title: "4.7 practice"
output: html_notebook
---

```{r}
library(rethinking)
```

### 4M1. For the model definition below, simulate observed y values from the prior (not the posterior).

$$
y_i \sim Normal(μ, σ) \\
μ \sim Normal(0, 10) \\
σ \sim Exponential(1) \\
$$

```{r}
mu <- rnorm(n = 1e4, mean = 0, sd = 10)
sigma <- rexp(n = 1e4, rate = 1)
y = rnorm(n = 1e4, mean = mu, sd = sigma)
dens(y)
```

### 4M2. Translate the model just above into a quap formula

```{r}
model <- quap(
    alist(
        sigma ~ dexp(1),
        mu ~ dnorm(mean=0, sd=10),
        y ~ dnorm(mean = mu, sd = sigma)
    ), data = data.frame(runif(10))
)
```

### 4M3. Translate the quap model formula below into a mathematical model definition.

```{r}
flist <- alist(
    y ~ dnorm( mu , sigma ),
    mu <- a + b*x,
    a ~ dnorm( 0 , 10 ),
    b ~ dunif( 0 , 1 ),
    sigma ~ dexp( 1 )
)
```

$$
y_i \sim Normal(\mu_i, \sigma)\\
\mu_i = a + bx_i\\
a \sim Norm(0, 10)\\
b \sim Uniform(0, 1)\\
\sigma \sim Exponential(1)
$$

### 4M4. A sample of students is measured for height each year for 3 years. After the third year, you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.

Standardise heights to mean of zero, variance of 1.

- $h_i$ height of individual $i$
- $y$ year
- $a_i$ per individual intercept
- $b$ growth rate

$$
h_{iy} \sim Normal(\mu, \sigma)\\
\mu = a_i + by\\
a \sim Normal(0, 1)\\
b \sim Normal(0, 1)\\
\sigma \sim Exponential(1)
$$

- Could restrict prior on b to favour positive growth rates.
- Could have individual growth rates
- Could have hierarchical model, but I don't think we've got there yet...

### 4M5. Now suppose I remind you that every student got taller each year. Does this information lead you to change your choice of priors? How?

$$
b \sim HalfNormal(1)
$$

### 4M6. Now suppose I tell you that the variance among heights for students of the same age is never more than 64cm. How does this lead you to revise your priors?

Could restrict $\sigma$ even more, by giving values of $\sigma$ that correspond to >64cm 0 prior probability.


# Hard


### 4H1. The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals for each of these individuals. That is, fill in the table below, using model-based predictions.


```{r}
# Weights of individuals with unknown heights
weights <- c(46.95, 43.72, 64.78, 32.59, 54.63)

# Load data
data("Howell1")
d <- Howell1

# Standardise weight
d$weight_s <- (d$weight - mean(d$weight)) / sd(d$weight)

# Make weight squared variable
d$weight_s2 <- d$weight_s ^ 2

# Model lifted from page 111, box 4.65
m4.5 <- quap(
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + b1 * weight_s + b2 * weight_s2,
        a ~ dnorm(178, 20),
        b1 ~ dlnorm(0, 1),
        b2 ~ dnorm(0, 1),
        sigma ~ dunif(0, 50)
    ), data = d
)

precis(m4.5)
```
Plot the model fit (for completeness)

```{r}
weight_seq = seq(-2.2, 2, length.out = 30)
pred_dat <- list(weight_s=weight_seq, weight_s2=weight_seq^2 )
mu <- link(m4.5, data=pred_dat)
mu_mean <- apply(mu, 2, mean)
mu_pi <- apply(mu, 2, PI, prob=0.89)
sim_height <- sim(m4.5, data = pred_dat)
height_pi <- apply(sim_height, 2, PI, prob=0.89)

plot(height ~ weight_s, d, col=col.alpha(rangi2, 0.5))
lines(weight_seq, mu_mean)
shade(mu_pi, weight_seq)
shade(height_pi, weight_seq)
```

Now, back to the actual question... Predict the heights for the 5 individuals.

```{r}
unknown = data.frame(weights)
unknown$weights_s <- (unknown$weights - mean(d$weight)) / sd(d$weight)
unknown$weights_s2 <- unknown$weights_s ^ 2

mu = link(m4.5, data=list(weight_s=unknown$weights_s, weight_s2=unknown$weights_s2))
unknown$mu_mean <- apply(mu, 2, mean)
unknown <- cbind(unknown, t(apply(mu, 2, HPDI, prob=0.89)))
unknown
```

### 4H2. Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it.

```{r}
d <- Howell1[Howell1$age < 18, ]
```

(a) Fit a linear regression to these data, using quap . Present and interpret the estimates. For
every 10 units of increase in weight, how much taller does the model predict a child gets?

```{r}
d$weight_s <- (d$weight - mean(d$weight)) / sd(d$weight)

lr <- quap(
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + b * weight_s,
        a ~ dnorm(mean(d$height), sd(d$height)),
        b ~ dnorm(0, 5),
        sigma ~ dexp(1)
    ), data=d
)

precis(lr)
```

Individual with mean weight has height of 108.32 cm.

Individual that is 10 units of weight heavier should be this much taller:

```{r}
w <- 10 / sd(d$weight)  # Standardize '10 units of weight heavier' 
diff <- link(lr, data = list(weight_s = c(0, w)))
mean(diff[,2] - diff[,1])

# Could also have done:
23.97 * w
```

(b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Super-
impose the MAP regression line and 89% interval for the mean. Also superimpose the 89% interval
for predicted heights.

```{r}
weight_seq <- seq(-2, 3, length.out = 30)
mu <- link(lr, data = list(weight_s=weight_seq))
mu_mean <- apply(mu, 2, mean)
mu_hpdi <- apply(mu, 2, HPDI, prob=0.89)
sim_height <- sim(lr, data=list(weight_s=weight_seq))
sim_height_pi <- apply(sim_height, 2, HPDI, prob=0.89)

plot(height ~ weight_s, d, col=col.alpha(rangi2,0.5))
lines(weight_seq, mu_mean)
shade(mu_hpdi, weight_seq)
shade(sim_height_pi, weight_seq)
```

(c) What aspects of the model fit concern you? Describe the kinds of assumptions you would
change, if any, to improve the model. You don’t have to write any new code. Just explain what the
model appears to be doing a bad job of, and what you hypothesize would be a better model.

- Doesn't capture curve in data.
- Posterior predictive looks quite dissimilar to the actual data.
- Could use some kind of curve / spline.
