---
title: "3.5 practice"
output: html_notebook
---

```{r message=FALSE}
library(rethinking)
```

## Easy

Generate some samples to work with:

```{r}
# Copy pasted from book
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid, posterior, type = "l")
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

### 3E1. How much posterior probability lies below p = 0.2?

```{r}
sum(samples < 0.2) / length(samples)
```
### 3E2. How much posterior probability lies above p = 0.8?

```{r}
sum(samples > 0.8) / length(samples)
```


### 3E3. How much posterior probability lies between p = 0.2 and p = 0.8?

```{r}
sum((samples < 0.8) & (samples > 0.2)) / length(samples)
```

### 3E4. 20% of the posterior probability lies below which value of p?

```{r}
quantile(samples, 0.2)
```


### 3E5. 20% of the posterior probability lies above which value of p?

```{r}
quantile(samples, (1 - 0.2))
```

### 3E6. Which values of p contain the narrowest interval equal to 66% of the posterior probability?

Highest posterior density interval

```{r}
HPDI(samples, prob = 0.66)
```


### 3E7. Which values of p contain 66% of the posterior probability, assuming _equal posterior probability both below and above the interval_?

```{r}
PI(samples, prob = 0.66)
```

---

## Medium

### 3M1. Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.

```{r}
p_grid = seq(0, 1, length.out = 1e4)
prior = rep(1, 1e4)
lik = dbinom(8, 15, prob = p_grid)
posterior = lik * prior
posterior = posterior / sum(posterior)
plot(p_grid, posterior, type="l")
```

### 3M2. Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p.

```{r}
samples = sample(p_grid, size = 1e4, replace = TRUE, prob = posterior)
HPDI(samples, prob = 0.9)
```

### 3M3. Construct a posterior predictive check for this model and data. This means simulate the distribution of samples, averaging over the posterior uncertainty in p. What is the probability of observing 8 water in 15 tosses?

```{r}
posterior_predictive = rbinom(n = 1e4, size = 15, prob = samples)
simplehist(posterior_predictive, xlab="n. waters")
```
Probability of observing 8 waters:

```{r}
sum(posterior_predictive == 8) / length(posterior_predictive)
```

### 3M4. Using the posterior distribution constructed from the new (8/15) data, now calculate the probability of observing 6 water in 9 tosses.

```{r}
posterior_predictive = rbinom(1e4, size = 9, prob = samples)
sum(posterior_predictive == 6) / length(posterior_predictive)
```

### 3M5. Start over at 3M1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5. This corresponds to prior information that a majority of the Earth’s surface is water. Repeat each problem above and compare the inferences. What difference does the better prior make? If it helps, compare inferences (using both priors) to the true value p = 0.7.

```{r}
step_prior = ifelse(p_grid > 0.5, 1, 0)
plot(p_grid, step_prior, type = "l", main = "Step prior")
```

```{r}
likelihood = dbinom(x = 8, size = 15, prob = p_grid)
posterior = likelihood * step_prior
plot(p_grid, posterior, type = "l", main = "Posterior")
```
90% HPDI:

```{r}
samples = sample(p_grid, size = 1e4, replace = TRUE, prob = posterior)
HPDI(samples, 0.9)
```

Posterior predictive check. What is the probability of observing 8 waters in 15 tosses?

```{r}
posterior_predictive = rbinom(n = 1e4, size = 15, p = samples)
sum(posterior_predictive == 8) / length(posterior_predictive)
```

Probability of observing 6 water in 9 tosses:

```{r}
posterior_predictive = rbinom(n = 1e4, size = 9, p = samples)
sum(posterior_predictive == 6) / length(posterior_predictive)
```

What difference does the better prior make?

- More posterior probability around true value of p.

### 3M6. Suppose you want to estimate the Earth’s proportion of water very precisely. Specifically, you want the 99% percentile interval of the posterior distribution of p to be only 0.05 wide. This means the distance between the upper and lower bound of the interval should be 0.05. How many times will you have to toss the globe to do this?

Not 100% on this... What I've done assumes that 8/15 of the throws are water.  This would suggest ~2500 throws would be required for this accuracy.

```{r}
width_99_pi_interval <- function (throws, proportion = 8/15) {
    x = round(throws * proportion)
    likelihood = dbinom(x = x, size = throws, prob = p_grid)
    posterior = likelihood * step_prior
    samples = sample(p_grid, size = 1e5, replace = TRUE, prob = posterior)
    pi <- PI(samples, prob = 0.99)
    pi[2] - pi[1]
}

throws = c(50, 100, 250, 500, 1000, 2500, 5000)
width_99pi = sapply(throws, width_99_pi_interval)

plot(throws, width_99pi, type = "b")
```

```{r}
width_99_pi_interval(2500)
```

## Hard

### Introduction. The practice problems here all use the data below. These data indicate the gender (male=1, female=0) of officially reported first and second born children in 100 two-child families.

```{r}
data(homeworkch3)
head(birth1)
head(birth2)
sum(birth1) + sum(birth2)
```

### 3H1. Using grid approximation, compute the posterior distribution for the probability of a birth being a boy. Assume a uniform prior probability. Which parameter value maximizes the posterior probability?

```{r}
p_grid = seq(0, 1, length.out = 1e4)
likelihood = dbinom(x = sum(birth1) + sum(birth2),
                    size = length(birth1) + length(birth2),
                    prob = p_grid)
prior = rep(1, length(p_grid))
posterior = prior * likelihood
posterior = posterior / sum(posterior)
plot(p_grid, posterior, type="l")
```

P with maximum posterior probability:

```{r}
p_grid[which.max(posterior)]
```

### 3H2. Using the sample function, draw 10,000 random parameter values from the posterior distribution you calculated above. Use these samples to estimate the 50%, 89%, and 97% highest posterior density intervals.

```{r}
samples <- sample(x = p_grid, size=1e4, prob=posterior, replace = TRUE)
HPDI(samples, 0.5)
HPDI(samples, 0.89)
HPDI(samples, 0.97)
```

### 3H3. Use rbinom to simulate 10,000 replicates of 200 births. You should end up with 10,000 numbers, each one a count of boys out of 200 births. Compare the distribution of predicted numbers of boys to the actual count in the data (111 boys out of 200 births). There are many good ways to visualize the simulations, but the dens command (part of the rethinking package) is probably the easiest way in this case.

```{r}
posterior_predictive = rbinom(n = 1e4, size = 200, prob = samples)
dens(x = posterior_predictive, show.HPDI = 0.89)
```

Does it look like the model fits the data well? That is, does the distribution of predictions include the actual observation as a central, likely outcome?

- Yes, centre of the density is at about 111.

### 3H4. Now compare 10,000 counts of boys from 100 simulated first borns only to the number of boys in the first births, birth1. How does the model look in this light?

```{r}
posterior_predictive = rbinom(n = 1e4, size = 100, prob = samples)
dens(posterior_predictive, show.HPDI = 0.89)
```

Number of boys in first borns:

```{r}
sum(birth1)
```

- Model fit is worse - centre of the distribution is around 55-57, whereas the actual number of boys in first borns is 51.


### 3H5. The model assumes that sex of first and second births are independent. To check this assumption, focus now on _second births that followed female first borns_. Compare 10,000 simulated counts of boys to only those second births that followed girls. To do this correctly, you need to count the number of first borns who were girls and simulate that many births, 10,000 times. Compare the counts of boys in your simulations to the actual observed count of boys following girls. How does the model look in this light? Any guesses what is going on in these data?

```{r}
# Second births that followed female first borns
birth2_after_female = birth2[birth1 == 0]

# Number of births that followed a female first born
length(birth2_after_female)

# Number of males in that set
sum(birth2_after_female)
```

```{r}
posterior_predictive = rbinom(1e4, size = length(birth2_after_female), prob = samples)
simplehist(posterior_predictive)
```

Central values from the posterior prediction are ~26-29, whereas the actual number of males following females is 39, far higher.  Furthermore, only 1/4 of first births were female.

